---
title: "R Notebook"
author: "Jesid"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(here)
library(readxl)
library(dplyr)
library(ggplot2)
```
# Read Data
```{r}
df <- read_excel(here("data", "HomicideData_Original_revised.xlsx"), sheet="HomicideData")
colSums(is.na(df))
```

# Distribution of DV: All Years All States
```{r}
ggplot(data=df, aes(x=(NumHomicides/Population) *100000)) +
  geom_histogram(aes(y=..density..), col='black', fill='white') +
  geom_density(alpha=.6, fill="pink") +
  labs(title = "Distribution of Homicides: All States & Years")
```

# Preparing Data for Modeling
```{r}
# Removing DC - not a state
df <- df %>% filter(StateCode !="DC")
# Grouping Firearms
df <- df %>% mutate(Firearms = AnyOtherWeapon1 + Machinegun3 + ShortBarreledRifle5 + ShortBarreledShotgun6)
# Grouping Violent Crime Except Murder
df <- df %>% mutate(VCNotMurder = RapeVC + AggravatedAssaultVC + RobberyVC)
# Selecting data of interest for modeling
data <- df %>% select(Year, State, NumHomicides, Firearms, NumDrugUsers, NumAlcoholUsers, SeriousMentalIllness,
                      Population, RealGDP, GiniIndex, UnenemploymentRate, LawEnforcementOfficers, 
                      CannabisMedical, CannabisRecreational, SubAbuseInpatientCareBeds,
                      VCNotMurder,PropertyCrime, White, Black, Hispanic, Asian, `Children 0-18`, `Adults 19-25`, `Adults 26-34`, Education,
                      PermitAllFirearms, reportall, defactoreg, relinquishment, UniversalBackgroundChecksAllFirearms, UniversalPermit
                      )
# Rename Variables in Data
data <- data %>% rename(ViolentCrime = VCNotMurder, BedsForAbuse = SubAbuseInpatientCareBeds,
                        UnemploymentRate = UnenemploymentRate,
                        ReportAllFirearms = reportall, GunRegistration = defactoreg,
                        Children18orUnder = `Children 0-18`,
                        RelinquishGun = relinquishment,
                        Adult19to25 = `Adults 19-25`,
                        Adult26to34 = `Adults 26-34`,
                        UniversalBackgroundCheck = UniversalBackgroundChecksAllFirearms)
# Checking data
colSums(is.na(data))
```

## Creating New Variable Transformation
```{r}
# Factorize
cols <- c("PermitAllFirearms", "ReportAllFirearms", "GunRegistration", "CannabisMedical", "CannabisRecreational","RelinquishGun", "UniversalBackgroundCheck")
for (col in cols){
  data[[col]] <- factor(data[[col]], levels = c(0,1), labels = c("No", "Yes"))  
}

# Year to Factor  
data$Year <- relevel(factor(data$Year), "2010")

# convert to Pop estimate - all except laws, gini, and unemployment rate
unit_of_people <- 100000
# cols
cols <- c("NumDrugUsers",  "NumAlcoholUsers", "LawEnforcementOfficers", "SeriousMentalIllness", "ViolentCrime", 
          "PropertyCrime", "BedsForAbuse", "NumHomicides", "Firearms", "White", "Black", "Hispanic", "Asian",
          "Children18orUnder","Adult19to25", "Adult26to34")
data[cols] <- (data[cols] / data[["Population"]]) * unit_of_people
# Real GDP per Capita
data$RealGDP <- data$RealGDP/data$Population
# convert Gini to percentage
data['GiniIndex'] = data['GiniIndex'] * 100
```

## Create Lags
```{r}
# testing lags
data <- data %>%
  arrange(Year) %>%
  group_by(State) %>%
  mutate(OneLagLawEnforcement = lag(LawEnforcementOfficers),
         OneLagGunPermit = lag(PermitAllFirearms),
         OneLagGunRegister = lag(GunRegistration),
         OneLagGunReport = lag(ReportAllFirearms),
         OneLagGunRelinquish = lag(RelinquishGun),
         OneLagBackground = lag(UniversalBackgroundCheck),
         OneLagCannabisRec = lag(CannabisRecreational), 
         OneLagCannabisMed = lag(CannabisMedical),
         OneLagViolentCrime = lag(ViolentCrime),
         OneLagFirearms = lag(Firearms),
         TwoLagLawEnforcement = lag(LawEnforcementOfficers, 2),
         TwoLagGunPermit = lag(PermitAllFirearms, 2),
         TwoLagGunRegister = lag(GunRegistration, 2),
         TwoLagGunReport = lag(ReportAllFirearms, 2),
         TwoLagGunRelinquish = lag(RelinquishGun, 2),
         TwoLagBackground = lag(UniversalBackgroundCheck, 2),
         TwoLagCannabisRec = lag(CannabisRecreational, 2), 
         TwoLagCannabisMed = lag(CannabisMedical, 2),
         TwoLagViolentCrime = lag(ViolentCrime, 2),
         TwoLagFirearms = lag(Firearms,2))

```


# Modeling
## Fixed Effects - Causes
```{r}
# base model - only causes
lm1 <- glm(log(NumHomicides) ~ log(Firearms) + NumDrugUsers + NumAlcoholUsers + SeriousMentalIllness +
           GiniIndex + UnemploymentRate + log(RealGDP)  + Black + Hispanic + Asian + 
           Children18orUnder + Adult19to25 + Education + Year + State, 
           data=data, family = gaussian(link = "identity"))

plot(lm1, which = c(1,2))
# Normality
shapiro.test(lm1$res)
# Homoskedasticity
bartlett.test(list(lm1$res, lm1$fit))
library(car)
vif(lm1)
library(lmtest)
dwtest(lm1)
# stargazer::stargazer(lm1, type="text", single.row = T)
```
## Fixed Effects - Causes & Interventions(Law Enforcement)
```{r}
# base model - only causes
lm2 <- glm(log(NumHomicides) ~ log(Firearms) + NumDrugUsers + NumAlcoholUsers + SeriousMentalIllness +
           GiniIndex + UnemploymentRate + log(RealGDP) + log(LawEnforcementOfficers) + Black + Hispanic + 
             Asian + Children18orUnder + Adult19to25 + Education + Year + State, 
           data=data, family = gaussian(link = "identity"))

plot(lm2, which = c(1,2))
# Normality
shapiro.test(lm2$res)
# Homoskedasticity
bartlett.test(list(lm2$res, lm2$fit))
library(car)
vif(lm2)
library(lmtest)
dwtest(lm2)
# stargazer::stargazer(lm1, lm2, type="text", single.row = T)
```
## Fixed Effects - With Policies
```{r}
# base model - only causes
lm3 <- glm(log(NumHomicides) ~ log(Firearms) + (PermitAllFirearms + GunRegistration + RelinquishGun + UniversalBackgroundCheck) + ReportAllFirearms + 
             NumDrugUsers + NumAlcoholUsers + SeriousMentalIllness + GiniIndex + UnemploymentRate  +
             log(RealGDP) + log(LawEnforcementOfficers) + Black + Hispanic + 
             Asian + Children18orUnder + Adult19to25 + Education + Year + State, 
           data=data, family = gaussian(link = "identity"))

plot(lm3, which = c(1,2))
# Normality
shapiro.test(lm3$res)
# Homoskedasticity
bartlett.test(list(lm3$res, lm3$fit))
library(car)
vif(lm3)
library(lmtest)
dwtest(lm3)
# stargazer::stargazer(lm1, lm2, lm3, type="text", single.row = T)
```
## Fixed Effects - With Policies & Interactions
```{r}
# base model - only causes
lm4 <- glm(log(NumHomicides) ~ log(Firearms)*(PermitAllFirearms + UniversalBackgroundCheck) + GunRegistration + RelinquishGun + ReportAllFirearms + 
             NumDrugUsers + NumAlcoholUsers + SeriousMentalIllness + GiniIndex + UnemploymentRate + 
             log(RealGDP) + log(LawEnforcementOfficers) + Black + Hispanic + 
             Asian + Children18orUnder + Adult19to25 + Education + Year + State, 
           data=data, family = gaussian(link = "identity"))

plot(lm4, which = c(1,2))
# Normality
shapiro.test(lm4$res)
# Homoskedasticity
bartlett.test(list(lm4$res, lm4$fit))
library(car)
vif(lm4)
library(lmtest)
dwtest(lm4)
```

### Compare Linear Models
```{r}
stargazer::stargazer(lm1, lm2, lm3, lm4, type="text", single.row = T)
```

## Fixed Effects - Lags On Laws
```{r}
# base model - only causes
lm5 <- glm(log(NumHomicides) ~ log(Firearms)*(OneLagGunPermit + OneLagBackground ) + OneLagGunRegister + OneLagGunRelinquish + OneLagGunReport + 
             NumDrugUsers + NumAlcoholUsers + SeriousMentalIllness + GiniIndex + UnemploymentRate + 
             log(RealGDP) + log(OneLagLawEnforcement) + Black + Hispanic + 
             Asian + Children18orUnder + Adult19to25 + Education + Year + State, 
           data=data, family = gaussian(link = "identity"))

plot(lm5, which = c(1,2))
# Normality
shapiro.test(lm5$res)
# Homoskedasticity
bartlett.test(list(lm5$res, lm5$fit))
library(car)
vif(lm5)
library(lmtest)
dwtest(lm5)
```
## Fixed Effects - Lags On Laws (Using Poisson Distribution)
```{r}
# base model - only causes
lm6 <- glm(round(NumHomicides,0) ~ log(Firearms)*(OneLagGunPermit + OneLagBackground) + OneLagGunRegister + OneLagGunRelinquish + OneLagGunReport + 
             NumDrugUsers + NumAlcoholUsers + SeriousMentalIllness + GiniIndex + UnemploymentRate + 
             log(RealGDP) + log(OneLagLawEnforcement) + Black + Hispanic + 
             Asian + Children18orUnder + Adult19to25 + Education + Year + State, 
           data=data, family = poisson(link = "log"))

plot(lm6, which = c(1,2))

# Homoskedasticity
bartlett.test(list(lm6$res, lm6$fit))
library(car)
# Multicollineairty
vif(lm6)
# Independence
dwtest(lm6)
```

```{r}
stargazer::stargazer(lm5, lm6, type="text", single.row = T)
```



## Model with Multilevels - Causes & Interventions(Without Lags)
```{r}
library(lme4)

re1 <- lmer(log(NumHomicides) ~ NumDrugUsers + NumAlcoholUsers + SeriousMentalIllness +
            GiniIndex + UnemploymentRate + log(RealGDP) + log(LawEnforcementOfficers) +
            log(Firearms) + (PermitAllFirearms + RelinquishGun + GunRegistration) + ReportAllFirearms +
            OneLagCannabisRec + OneLagCannabisMed + (1| State) +  (1 | Year) + Black + Hispanic + Asian + 
            Children18orUnder  + Education, 
            data=data, REML = F)
# ranef(re1)

# Normality - Failing
shapiro.test(residuals(re1))

# Homoskedasticity - Failing
bartlett.test(list(residuals(re1), fitted(re1)))

# Multicollinearity - Failing
vif(re1)
```



## Assumption Checking LMM
```{r}
library(DHARMa)
simulationOutput <- simulateResiduals(fittedModel = re1, plot = F)
plot(simulationOutput)
```
## Model with Multilevels - Causes & Interventions(With Interactions)
```{r}
library(lme4)

re2 <- lmer(log(NumHomicides) ~ NumDrugUsers*CannabisRecreational + NumAlcoholUsers + SeriousMentalIllness +
            GiniIndex + UnemploymentRate + log(RealGDP) + log(LawEnforcementOfficers) +
            log(Firearms)*(PermitAllFirearms) + GunRegistration + RelinquishGun + ReportAllFirearms +
            CannabisMedical + (1| State) +  (1 | Year) + Black + Hispanic + Asian + 
            Children18orUnder + Education, data=data)
# ranef(re2)

# Multicollinearity - Failing
vif(re2)
```

```{r}
# Assumption Checking
simulationOutput <- simulateResiduals(fittedModel = re2, plot = F)
plot(simulationOutput)
testDispersion(simulationOutput = simulationOutput)
```


## Model with Multilevels - Causes & Interventions (With Lags)
```{r}
re3 <- lmer(log(NumHomicides) ~ NumDrugUsers + NumAlcoholUsers + SeriousMentalIllness +
            GiniIndex + UnemploymentRate + log(RealGDP) + log(OneLagLawEnforcement) +
            log(Firearms) + (OneLagGunRegister+OneLagGunPermit) + RelinquishGun + OneLagGunReport  + OneLagCannabisRec + 
            (1| State) +  (1 | Year) + Black + Hispanic + Asian + Children18orUnder + Education, 
            data=data, REML = F)
# ranef(re3)
```

## Model with Multilevels - Causes & Interventions (With Lags & Interaction)
```{r}
re4 <- lmer(log(NumHomicides) ~ NumDrugUsers + CannabisRecreational + NumAlcoholUsers + SeriousMentalIllness +
            GiniIndex + UnemploymentRate + log(RealGDP) + log(LawEnforcementOfficers) +
            log(Firearms) + (PermitAllFirearms) + UniversalBackgroundCheck + GunRegistration + RelinquishGun + ViolentCrime + ReportAllFirearms
            + (1| State) +  (1 | Year) + Black + Hispanic + Asian + Children18orUnder + 
            Education, data=data, REML = F)
# ranef(re4)

# Multicollinearity - Failing
vif(re4)
```

## Check Assumptions
```{r}
# Assumption Checking
simulationOutput <- simulateResiduals(fittedModel = re4, plot = F)
plot(simulationOutput)
testDispersion(simulationOutput = simulationOutput)
```
### Compare Poisson & Log
```{r, echo=FALSE, warning=FALSE}
stargazer::stargazer( re1, re2, re3, re4, type="text", single.row = T, column.labels = c("Base Model", "Interactions", "Lagged Effects", "Lag&Interaction"))
```


## Model with Multilevels - Adding VC Interaction with Relinquish
```{r}
library(lme4)

re5 <- lmer(log(NumHomicides) ~ NumDrugUsers*CannabisRecreational + NumAlcoholUsers + SeriousMentalIllness +
            GiniIndex + UnemploymentRate + log(RealGDP) + log(LawEnforcementOfficers) +
            log(Firearms)*(PermitAllFirearms) + UniversalBackgroundCheck + GunRegistration + RelinquishGun*ViolentCrime + ReportAllFirearms
            + (1| State) +  (1 | Year) + Black + Hispanic + Asian + Children18orUnder + 
            Education, data=data, REML = F)
# ranef(re1)

# Multicollinearity - Failing
vif(re5)
# Assumption Checking
simulationOutput <- simulateResiduals(fittedModel = re5, plot = F)
plot(simulationOutput)
testDispersion(simulationOutput = simulationOutput)
```
## Model with Multilevels - Adding Multiple Lags
```{r}
re6 <- lmer(log(NumHomicides) ~ NumDrugUsers*(OneLagCannabisRec + CannabisRecreational) + NumAlcoholUsers + SeriousMentalIllness +
            GiniIndex + UnemploymentRate + log(RealGDP) + log(OneLagLawEnforcement) + log(LawEnforcementOfficers) +
            (log(OneLagFirearms) + log(Firearms))*(OneLagGunPermit + PermitAllFirearms) + OneLagBackground + UniversalBackgroundCheck + OneLagGunRegister + GunRegistration +
            RelinquishGun*(ViolentCrime + OneLagViolentCrime) + OneLagGunReport + ReportAllFirearms
            + (1| State) +  (1 | Year) + Black + Hispanic + Asian + Children18orUnder + 
            Education, data=data, REML = F)
# ranef(re1)

# Multicollinearity - Failing
vif(re6)
# Assumption Checking
simulationOutput <- simulateResiduals(fittedModel = re6, plot = F)
plot(simulationOutput)
testDispersion(simulationOutput = simulationOutput)
```
## Model with Multilevels - 2 Lags with Multiple
```{r}
re7 <- lmer(log(NumHomicides) ~ NumDrugUsers*(OneLagCannabisRec + CannabisRecreational) + NumAlcoholUsers + SeriousMentalIllness +
            GiniIndex + UnemploymentRate + log(RealGDP) + log(TwoLagLawEnforcement) + log(OneLagLawEnforcement) + log(LawEnforcementOfficers) +
            (log(OneLagFirearms) + log(Firearms))*(OneLagGunPermit + PermitAllFirearms) + OneLagBackground + UniversalBackgroundCheck 
            + OneLagGunRegister + GunRegistration + RelinquishGun*(ViolentCrime + OneLagViolentCrime)
            + OneLagGunReport + ReportAllFirearms 
            + (1| State) +  (1 | Year) + Black + Hispanic + Asian + Children18orUnder + 
            Education, data=data, REML = F)
# ranef(re7)

# Multicollinearity - Failing
vif(re7)
# Assumption Checking
simulationOutput <- simulateResiduals(fittedModel = re7, plot = F)
plot(simulationOutput)
testDispersion(simulationOutput = simulationOutput)
```


### Compare VC Interaction + Additional Lags
```{r, echo=FALSE, warning=FALSE}
stargazer::stargazer(re5, re6, re7, type="text", single.row = T, column.labels = c("Interactions", "OneLag", "TwoLag"))
# try logs on populatison
```
# Corr Plot
```{r}
temp <- data %>% select(NumHomicides, ViolentCrime, NumDrugUsers, NumAlcoholUsers, SeriousMentalIllness, GiniIndex,
                        UnemploymentRate,RealGDP,LawEnforcementOfficers,Firearms,Black,Hispanic,Asian,Children18orUnder,Education)
temp <- subset(temp, select = -State)
PerformanceAnalytics::chart.Correlation(temp, pch= "+")
correlations=cor(temp,use = "pairwise.complete.obs")
corrplot::corrplot(correlations,method = "number",type="upper")
```


## Effects
```{r}
library(effects)
effects <- allEffects(re5)
#Plot All
for (i in 1:length(effects)) {
  print(plot(effects, selection = i))
}
```

